---
title: "T32/F32: Research-In-Progress"
author: "Anish Sanjay Shah, MD/MS<br><span style = 'font-size: 70%;'>Cardiology Fellow<br>University of Illinois at Chicago</span>"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css"]
    self_contained: false
    nature:
      titleSlideClass: [left, middle, inverse]
      highlightLines: true
---

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)

print.tbl_df <- function(x, ...) {
	print.data.frame(x, right = FALSE, ...)
	invisible(x)
}

# Options
options(digits = 2)
```

```{r themer, echo=FALSE, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
	base_color = "#374e55", # gray
	colors = c(
		emory = "#00a1d5", # blue
		uic = "#df8f44", # orange
		computer = "#79af97", # green
		clinical = "#b24745", # red
		purple = "#6a6599", # purple
		brown = "80796b" # brown
	),
	footnote_position_bottom = "1px",
	extra_css = list(
		".remark-slide-number" = list("display" = "none")
	)
)
```

# .center[Weekly Updates]

---

class: inverse, middle
# July 11, 2022

---
class: inverse

.pull-left[

.emory[Mental Stress and Autonomic Reactivity] 
- Result section being completed
- Table and figures complete
- _Share with coauthors_

.emory[HRV and Coronary Vascular Tone]
- Enrolling new patients with continuous ECG recordings 
- Revisitng analyses

.clinical[Endocarditis]
- _Outlining_ review paper on arrhythmias and IE 
- Pixel density being _analyzed_ for vegetation chronicity
- IRB pending
]

.pull-right[
.uic[AF Registry]
- _Revising_ data dictionary
- Obtaining EGM data from EP stations at UIC and JBVA

.uic[Isoproterenol washout and membrane potentials]
- Researching background data
- Evaluating potential test models (animal versus cellular)

.computer[Modeling software]
- _Revising_ R package on model formulas and DAGs

.computer[Circadian rhythms]
- Algorithm for extracting rhythm patterns from continuous data
- Cosinor analysis modification in R
]

.footnote[
.emory[Stress Epidemiology], .uic[Arrhythmia Mechanisms], .computer[Biostatistics & Signal Processing], .clinical[Clinical]
]

