---
title: "T32/F32: Research-In-Progress"
author: "Anish Sanjay Shah, MD/MS<br><span style = 'font-size: 70%;'>Cardiology Fellow<br>University of Illinois at Chicago</span>"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css"]
    self_contained: false
    nature:
      titleSlideClass: [left, middle, inverse]
      highlightLines: true
---

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)
library(gt)

print.tbl_df <- function(x, ...) {
	print.data.frame(x, right = FALSE, ...)
	invisible(x)
}

# Options
options(digits = 2)
```
	
```{r, knitr, include=FALSE}
knitr::opts_chunk$set(
   fig.retina = 3,
   warning = FALSE,
   message = FALSE,
   echo = FALSE,
   cache = TRUE
)
```

```{r themer, include=FALSE}
library(scales)
library(ggsci)

# Golf scoring, lower is darker, higher is lighter
stress <- pal_material("blue", reverse = TRUE)(9)
arrhythmia <- pal_material("yellow", reverse = TRUE)(9)
clinical <- pal_material("pink", reverse = TRUE)(9)
computational <- pal_material("teal", reverse = TRUE)(9)
text <- pal_material("blue-grey", reverse = TRUE)(9)

library(xaringanthemer)
style_mono_light(
	base_color = text[1],
	white_color = text[9],
	colors = c(
		stress = stress[5],
		rhythm = arrhythmia[2],
		clin = clinical[3],
		comp = computational[4],
		grey = text[5],
		light = text[7]
	),
	footnote_position_bottom = "1px",
	extra_css = list(
		".remark-slide-number" = list("display" = "none")
	)
)

```

```{r projects, echo=FALSE}
st <- "stress"
ar <- "arrhythmia"
cl <- "clinical"
co <- "computational"

# Starting point, to be updated each week in teh weekly report
updates <- tribble(
	~Type, ~Project, ~Status, ~Effort,
	st, "MS and CV Mortality", "draft to coauthors", .2,
	ar, "AF Registry", "data validation started, revisit variables", .2,
	cl, "PERT", "education and research IRB to submit", .05,
	ar, "F32/T32", "reviewing autonomic literature at translational/basic science level", 0.2,
	ar, "K23", "not yet started", 0,
	cl, "Studying", "general cardiology and echo boards", 0.05,
	cl, "UIC Clinic", "weekly clinic with Mary", 0.10,
	cl, "Endocarditis and Pixel Density", "need to process images still", 0.05,
	ar, "Endocarditis and Arrhythmias", "students working on outline", 0.01,
	co, "Modeling Software", "writing code for generation of forest plots by subgroups {R}", 0.02,
	co, "EGM Signal Processing", "practicing data extraction from UIC with OpenEP, need access to JBVA data exports still {MATLAB}", 0.05,
	st, "HRV and CAD", "research team is currently enrolling on my behalf", 0.01,
	cl, "EP Lab", "considering 1-2 cases per week", 0.10
)

```


# .center[Weekly Updates]

---

class: inverse, middle
# July 11, 2022

---
class: inverse

.pull-left[

.stress[Mental Stress and Autonomic Reactivity] 
- Result section being completed
- Table and figures complete
- _Share with coauthors_

.stress[HRV and Coronary Vascular Tone]
- Enrolling new patients with continuous ECG recordings 
- Revisitng analyses

.clin[Endocarditis]
- _Outlining_ review paper on arrhythmias and IE 
- Pixel density being _analyzed_ for vegetation chronicity
- IRB pending
]

.pull-right[
.rhythm[AF Registry]
- _Revising_ data dictionary
- Obtaining EGM data from EP stations at UIC and JBVA

.rhythm[Isoproterenol washout and membrane potentials]
- Researching background data
- Evaluating potential test models (animal versus cellular)

.comp[Modeling software]
- _Revising_ R package on model formulas and DAGs

.comp[Circadian rhythms]
- Algorithm for extracting rhythm patterns from continuous data
- Cosinor analysis modification in R
]

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]


---

class: inverse, middle
# July 18, 2022

---
class: inverse

```{r}
updates <-
	updates |>
	dplyr::arrange(desc(Effort)) 

updates |>
	gt(rowname_col = "Project") |>
	tab_stubhead("Projects") |>
	cols_hide("Type") |>
	fmt_percent(columns = "Effort", decimals = 0) |>
	tab_style(
		style = cell_text(align = "left"),
		locations = cells_stub()
	) |>
	cols_align(align = "right", columns = c("Effort")) |>
	tab_style(
		style = list(cell_fill(color = stress[5]),
								 cell_borders(sides = "all", color = stress[5])),
		locations = cells_stub(rows = Type == "stress")
	) |>
	tab_style(
		style = list(cell_fill(color = arrhythmia[2]),
								 cell_borders(sides = "all", color = arrhythmia[2])),
		locations = cells_stub(rows = Type == "arrhythmia")
	) |>
	tab_style(
		style = list(cell_fill(color = clinical[5]),
								 cell_borders(sides = "all", color = clinical[5])),
		locations = cells_stub(rows = Type == "clinical")
	) |>
	tab_style(
		style = list(cell_fill(color = computational[5]),
								 cell_borders(sides = "all", color = computational[5])),
		locations = cells_stub(rows = Type == "computational")
	) |>
	tab_style(
		style = list(cell_fill(color = text[1]), 
								 cell_text(color = "white"),
								 cell_borders(sides = "all", color = text[1])),
		locations = list(cells_stubhead(), cells_column_labels())
	) |>
	tab_style(
		style = list(cell_borders(sides = "left", 
															color = text[1], 
															weight = px(2))),
		locations = list(cells_body(columns = c("Effort")),
										 cells_stub())
	) |>
	tab_style(
		style = list(cell_borders(sides = "all", 
															color = text[1], 
															weight = px(2))),
		locations = list(cells_column_labels(columns = everything()),
										 cells_stubhead())
	) |>
	tab_style(
		style = cell_fill(color = text[9]),
		locations = cells_body(columns = c("Status", "Effort"))
	) |>
	tab_style(
		style = cell_borders(sides = "right", 
												 color = text[9]),
		locations = cells_body(columns = c("Effort"))
	) |>
	tab_style(
		style = cell_text(color = "white"),
		locations = list(cells_stub())
	)
	
```

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---

## AF Registry

The AF registry is described [here](https://asshah4.github.io/afib). Current issues:

1. Data quality
1. Missingness
1. Variable selection
1. Management
1. Adjudication/review

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]
