---
title: "T32/F32: Research-In-Progress"
author: "Anish Sanjay Shah, MD/MS<br><span style = 'font-size: 70%;'>Cardiology Fellow<br>University of Illinois at Chicago</span>"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "custom.css"]
    self_contained: false
    nature:
      titleSlideClass: [left, middle, inverse]
      highlightLines: true
---

<style type="text/css">
.remark-code {font-size: 50%}
.remark-inline-code {font-size: 70%}
</style>

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)

print.tbl_df <- function(x, ...) {
	print.data.frame(x, right = FALSE, ...)
	invisible(x)
}

# Options
options(digits = 2)
```
	
```{r, knitr, include=FALSE}
knitr::opts_chunk$set(
   fig.retina = 3,
   warning = FALSE,
   message = FALSE,
   echo = FALSE,
   cache = FALSE
)
```

```{r themer, include=FALSE}
library(scales)
library(ggsci)

# Golf scoring, lower is darker, higher is lighter
stress <- pal_material("blue", reverse = TRUE)(9)
arrhythmia <- pal_material("yellow", reverse = TRUE)(9)
clinical <- pal_material("pink", reverse = TRUE)(9)
computational <- pal_material("teal", reverse = TRUE)(9)
text <- pal_material("blue-grey", reverse = TRUE)(9)

library(xaringanthemer)
style_mono_accent(
	base_color = text[1],
	black_color = text[1],
	white_color = "white",
	link_color = computational[4],
	inverse_link_color = computational[4],
	colors = c(
		stress = stress[4],
		rhythm = arrhythmia[2],
		clin = clinical[3],
		comp = computational[4],
		grey = text[5],
		light = text[7]
	),
	footnote_position_bottom = "1px",
	extra_css = list(
		".remark-slide-number" = list("display" = "none")
	)
)

```

```{r projects, echo=FALSE}
st <- "stress"
ar <- "arrhythmia"
cl <- "clinical"
co <- "computational"

# Starting point, to be updated each week in teh weekly report
updates <- tribble(
	~Type, ~Project, ~Status, ~Effort,
	st, "MS and CV Mortality", NA_character_, 0,
	ar, "AF Registry", NA_character_, 0,
	cl, "PERT", NA_character_, 0,
	ar, "F32/T32/K23", NA_character_, 0, 
	cl, "Studying", NA_character_, 0,
	cl, "UIC Clinic", NA_character_, 0, 
	cl, "Endocarditis and Pixel Density", NA_character_, 0,
	ar, "Endocarditis and Arrhythmias", NA_character_, 0,
	co, "Modeling Software", NA_character_, 0,
	co, "EGM Signal Processing", NA_character_, 0,
	st, "HRV and CAD", NA_character_, 0,
	cl, "EP Lab", NA_character_, 0
)

# Thematic changes and management of columns and styles
library(gt)
tbl_research_updates <- function(data) {
	gt(data, rowname_col = "Project") |>
		tab_stubhead("Projects") |>
		cols_hide("Type") |>
		fmt_percent(columns = "Effort", decimals = 0) |>
		opt_table_lines(extent = "none") |>
		opt_row_striping(row_striping = FALSE) |>
		# Fill all borders to match background
		tab_style(
			style = cell_text(align = "left"),
			locations = cells_stub()
		) |>
		cols_align(align = "right", columns = c("Effort")) |>
		tab_style(
			style = list(cell_fill(color = stress[5]),
									 cell_borders(sides = "all", color = stress[5])),
			locations = cells_stub(rows = Type == "stress")
		) |>
		tab_style(
			style = list(cell_fill(color = arrhythmia[2]),
									 cell_borders(sides = "all", color = arrhythmia[2])),
			locations = cells_stub(rows = Type == "arrhythmia")
		) |>
		tab_style(
			style = list(cell_fill(color = clinical[5]),
									 cell_borders(sides = "all", color = clinical[5])),
			locations = cells_stub(rows = Type == "clinical")
		) |>
		tab_style(
			style = list(cell_fill(color = computational[5]),
									 cell_borders(sides = "all", color = computational[5])),
			locations = cells_stub(rows = Type == "computational")
		) |>
		tab_style(
			style = list(cell_fill(color = text[1]), 
									 cell_text(color = "white"),
									 cell_borders(sides = "all", color = text[1])),
			locations = list(cells_stubhead(), cells_column_labels())
		) |>
		tab_style(
			style = cell_borders(sides = "left", 
																color = text[1], 
																weight = px(2)),
			locations = cells_body(columns = c("Effort"))
		) |>
		tab_style(
			style = cell_borders(sides = "bottom", 
																color = text[1], 
																weight = px(2)),
			locations = list(cells_column_labels(columns = everything()),
											 cells_stubhead())
		) |>
		tab_style(
			style = list(cell_borders(sides = "all", 
																color = text[1], 
																weight = px(2))),
			locations = list(cells_column_labels(columns = everything()),
											 cells_stubhead())
		) |>
		tab_style(
			style = cell_fill(color = "white"),
			locations = cells_body(columns = everything())
		) |> 
		fmt_markdown(columns = everything()) |>
		cols_width(
			Project ~ pct(20),
			Status ~ pct(70),
			Effort ~ pct(10)
		) |>
		tab_style(style = cell_text(size = "x-small"),
							locations = cells_body(
								columns = Status
							))
}
```


# .center[Weekly Updates]

.pull-left[
- [July 11, 2022](#july-11-22)
- [July 18, 2022](#july-18-22)
- [July 25, 2022](#july-18-22)
]

---
class: inverse, middle
# July 11, 2022

---
class: inverse

.pull-left[

.stress[Mental Stress and Autonomic Reactivity] 
- Result section being completed
- Table and figures complete
- _Share with coauthors_

.stress[HRV and Coronary Vascular Tone]
- Enrolling new patients with continuous ECG recordings 
- Revisitng analyses

.clin[Endocarditis]
- _Outlining_ review paper on arrhythmias and IE 
- Pixel density being _analyzed_ for vegetation chronicity
- IRB pending
]

.pull-right[
.rhythm[AF Registry]
- _Revising_ data dictionary
- Obtaining EGM data from EP stations at UIC and JBVA

.rhythm[Isoproterenol washout and membrane potentials]
- Researching background data
- Evaluating potential test models (animal versus cellular)

.comp[Modeling software]
- _Revising_ R package on model formulas and DAGs

.comp[Circadian rhythms]
- Algorithm for extracting rhythm patterns from continuous data
- Cosinor analysis modification in R
]

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]


---
class: inverse, middle
# July 18, 2022

---
class: inverse, center

```{r}
updates <-
	updates |>
	mutate(Status = case_when(
		Project == "MS and CV Mortality" ~ 
			"Team = Amit Shah (*Emory*)  
			Draft to co-authors",
		Project == "AF Registry" ~ 
			"Data validation started, revisit variables",
		Project == "PERT" ~ 
			"Education and research IRB",
		Project == "F32/T32/K23" ~ 
			"Reviewing autonomic literature at translational/basic science level",
		Project == "Studying" ~ 
			"General cardiology/echo boards in **Fall 2023",
		Project == "UIC Clinic" ~ 
			"Weekly clinic with Mary",
		Project == "Endocarditis and Pixel Density" ~ 
			"Processing images pending",
		Project == "Endocarditis and Arrhythmias" ~ 
			"Outlining with student",
		Project == "Modeling Software" ~ 
			"Writing code for forest plots in {R}",
		Project == "EGM Signal Processing" ~ 
			"Practicing data extraction at UIC, but still need JBVA access {MATLAB}",
		Project == "HRV and CAD" ~ 
			"Emory Biobank continually recruiting",
		Project == "EP Lab" ~ 
			"Considering 1 to 2 cases this week",
		TRUE ~ Status
	)) |>
	mutate(Effort = case_when(
		Project == "MS and CV Mortality" ~ 0.2,
		Project == "AF Registry" ~ 0.2,
		Project == "PERT" ~ 0.05,
		Project == "F32/T32/K23" ~ 0.2,
		Project == "Studying" ~ 0.05,
		Project == "UIC Clinic" ~ 0.1,
		Project == "Endocarditis and Pixel Density" ~ 0.05,
		Project == "Endocarditis and Arrhythmias" ~ 0.02,
		Project == "Modeling Software" ~ 0.05,
		Project == "EGM Signal Processing" ~ 0.05,
		Project == "HRV and CAD" ~ 0.01,
		Project == "EP Lab" ~ 0.1,
		TRUE ~ Effort
	)) |>
	arrange(desc(Effort))

tbl_research_updates(updates)
```

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---

### AF Registry

The AF registry is described [here](https://asshah4.github.io/afib). Current issues:

1. Data quality - consistency between reviewers
1. Missingness - decisions on acceptable thresholds
1. Variable selection - echo findings, EP studies, medications, labs, symptoms
1. Management - REDCap, shared excel sheet
1. Adjudication/review - outcomes, clinical follow-up length

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---
class: inverse, middle
# July 25, 2022

---
class: inverse, center

```{r}
updates <-
	updates |>
	mutate(Status = case_when(
		Project == "MS and CV Mortality" ~ 
			"Team = Amit Shah (*Emory*)  
			Final figures of mortality by subgroups **7/25**  
			Revise discussion on vagal withdrawal **6/26**",
		Project == "AF Registry" ~ 
			"Decide on retrospective/prospective cohort definitions **TBD**",
		Project == "PERT" ~ 
			"Submit IRB for education and clinical data **7/28**",
		Project == "F32/T32/K23" ~ 
			"Aim 1 = Human vagolysis and SCD  
			**Aim 2 = Murine model of vagal withdrawal and arrhythmia**  
			Active research into model development **daily**",
		Project == "Studying" ~ 
			"General cardiology/echo boards in **Fall 2023",
		Project == "UIC Clinic" ~ 
			"Weekly clinic with Dr. Ziccardi",
		Project == "Endocarditis and Pixel Density" ~ 
			"Team = Lee Frazin, Satish Mishra, David Polman
			Revise still images to create standard for processing and SOP **7/30**",
		Project == "Endocarditis and Arrhythmias" ~ 
			"Meet with Atreya Mishra to complete outline **07/28**",
		Project == "Modeling Software" ~ 
			"Forest plots by subgroups in {R}, **drafted**",
		Project == "EGM Signal Processing" ~ 
			"Convert CARTO data into OpenEP **7/31**",
		Project == "HRV and CAD" ~ 
			"Emory Biobank continually recruiting  
			Meet with staff to assess data quality **TBD**",
		Project == "EP Lab" ~ 
			"Discuss with Adam which cases to cover this week  ",
		TRUE ~ Status
	)) |>
	mutate(Effort = case_when(
		Project == "MS and CV Mortality" ~ 0.2,
		Project == "AF Registry" ~ 0.1,
		Project == "PERT" ~ 0.1,
		Project == "F32/T32/K23" ~ 0.2,
		Project == "Studying" ~ 0.03,
		Project == "UIC Clinic" ~ 0.1,
		Project == "Endocarditis and Pixel Density" ~ 0.1,
		Project == "Endocarditis and Arrhythmias" ~ 0.02,
		Project == "Modeling Software" ~ 0.05,
		Project == "EGM Signal Processing" ~ 0.05,
		Project == "HRV and CAD" ~ 0.01,
		Project == "EP Lab" ~ 0.1,
		TRUE ~ Effort
	)) |>
	arrange(desc(Effort))
	
tbl_research_updates(updates)

```


.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---

### Forest Plots

.pull-left[
```{r, echo = TRUE}
# Forest Plots
library(volundr)
m <- rx(
	Petal.Length ~ X(Sepal.Length) + Petal.Width + S(Species),
	label = list(
		Petal.Length ~ "Length of Petals", 
		Species ~ "Genus of the Flower"
	),
	pattern = "direct"
	) |>
	fmls(order = 2) |>
	fit(.fit = lm,
			data = iris,
			archetype = TRUE) |>
	mdls()
```
]

.pull-right[
```{r, echo = TRUE}
tbl <- tbl_forest( #<<
	object = m,
	y = "Petal.Length",
	x = "Sepal.Length",
	groups = "Species",
	columns = list(
		beta ~ "Estimate",
		conf ~ "95% Confidence Interval",
		n ~ "Number of Samples"
	),
	axis = list(
		lim ~ c(0, 3),
		breaks ~ c(0, 1, 2),
		lab ~ "ß (95% CI)",
		title ~ "Petal Length by Sepal Length"
	)
)
```
]

A simple way to make sub-group analysis Forest plots, working on improving customization.

---

class: middle
```{r, echo = FALSE}
tbl |>
	tab_style(style = cell_fill(color = "white"),
						locations = cells_body(columns = everything()))
```


