---
title: "T32/F32: Research-In-Progress"
author: "Anish Sanjay Shah, MD/MS<br><span style = 'font-size: 70%;'>Cardiology Fellow<br>University of Illinois at Chicago</span>"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css"]
    self_contained: true
    anchor_sections: true
    nature:
      titleSlideClass: [left, middle, inverse]
      highlightLines: true
---

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)

print.tbl_df <- function(x, ...) {
	print.data.frame(x, right = FALSE, ...)
	invisible(x)
}

# Options
options(digits = 2)
```
	
```{r, knitr, include=FALSE}
knitr::opts_chunk$set(
   fig.retina = 3,
   warning = FALSE,
   message = FALSE,
   echo = FALSE,
   cache = TRUE
)
```

```{r themer, include=FALSE}
library(scales)
library(ggsci)

# Golf scoring, lower is darker, higher is lighter
stress <- pal_material("blue", reverse = TRUE)(9)
arrhythmia <- pal_material("yellow", reverse = TRUE)(9)
clinical <- pal_material("pink", reverse = TRUE)(9)
computational <- pal_material("teal", reverse = TRUE)(9)
text <- pal_material("blue-grey", reverse = TRUE)(9)

library(xaringanthemer)
style_mono_accent(
	base_color = text[1],
	black_color = text[1],
	white_color = "white",
	link_color = computational[4],
	inverse_link_color = computational[4],
	colors = c(
		stress = stress[4],
		rhythm = arrhythmia[2],
		clin = clinical[3],
		comp = computational[4],
		grey = text[5],
		light = text[7]
	),
	footnote_position_bottom = "1px",
	extra_css = list(
		".remark-slide-number" = list("display" = "none")
	)
)

```

```{r projects, echo=FALSE}
st <- "stress"
ar <- "arrhythmia"
cl <- "clinical"
co <- "computational"

# Starting point, to be updated each week in teh weekly report
updates <- tribble(
	~Type, ~Project, ~Status, ~Effort,
	st, "MS and CV Mortality", "draft to coauthors", .2,
	ar, "AF Registry", "data validation started, revisit variables", .2,
	cl, "PERT", "education and research IRB to submit", .05,
	ar, "F32/T32/K23", "reviewing autonomic literature at translational/basic science level", 0.2,
	cl, "Studying", "general cardiology and echo boards", 0.05,
	cl, "UIC Clinic", "weekly clinic with Mary", 0.10,
	cl, "Endocarditis and Pixel Density", "need to process images still", 0.05,
	ar, "Endocarditis and Arrhythmias", "students working on outline", 0.01,
	co, "Modeling Software", "writing code for generation of forest plots by subgroups {R}", 0.02,
	co, "EGM Signal Processing", "practicing data extraction from UIC with OpenEP, need access to JBVA data exports still {MATLAB}", 0.05,
	st, "HRV and CAD", "research team is currently enrolling on my behalf", 0.01,
	cl, "EP Lab", "considering 1-2 cases per week", 0.10
)

```

```{r gt, echo=FALSE}
library(gt)

# Thematic changes and management of columns and styles
tbl_research_updates <- function(data) {
	gt(data, rowname_col = "Project") |>
		tab_stubhead("Projects") |>
		cols_hide("Type") |>
		fmt_percent(columns = "Effort", decimals = 0) |>
		# Fill all borders to match background
		tab_style(
			style = cell_text(align = "left"),
			locations = cells_stub()
		) |>
		cols_align(align = "right", columns = c("Effort")) |>
		tab_style(
			style = list(cell_fill(color = stress[5]),
									 cell_borders(sides = "all", color = stress[5])),
			locations = cells_stub(rows = Type == "stress")
		) |>
		tab_style(
			style = list(cell_fill(color = arrhythmia[2]),
									 cell_borders(sides = "all", color = arrhythmia[2])),
			locations = cells_stub(rows = Type == "arrhythmia")
		) |>
		tab_style(
			style = list(cell_fill(color = clinical[5]),
									 cell_borders(sides = "all", color = clinical[5])),
			locations = cells_stub(rows = Type == "clinical")
		) |>
		tab_style(
			style = list(cell_fill(color = computational[5]),
									 cell_borders(sides = "all", color = computational[5])),
			locations = cells_stub(rows = Type == "computational")
		) |>
		tab_style(
			style = list(cell_fill(color = text[1]), 
									 cell_text(color = "white"),
									 cell_borders(sides = "all", color = text[1])),
			locations = list(cells_stubhead(), cells_column_labels())
		) |>
		tab_style(
			style = cell_borders(sides = "left", 
																color = text[1], 
																weight = px(2)),
			locations = cells_body(columns = c("Effort"))
		) |>
		tab_style(
			style = cell_borders(sides = "bottom", 
																color = text[1], 
																weight = px(2)),
			locations = list(cells_column_labels(columns = everything()),
											 cells_stubhead())
		) |>
		tab_style(
			style = list(cell_borders(sides = "all", 
																color = text[1], 
																weight = px(2))),
			locations = list(cells_column_labels(columns = everything()),
											 cells_stubhead())
		) |>
		tab_style(
			style = cell_fill(color = "white"),
			locations = cells_body(columns = everything())
		) |> 
		fmt_markdown(columns = everything()) |>
		cols_width(
			Project ~ pct(20),
			Status ~ pct(70),
			Effort ~ pct(10)
		) |>
		tab_style(style = cell_text(size = "x-small"),
							locations = "Status")
}
```


# .center[Weekly Updates]

.pull-left[
- [July 11, 2022](#july-11-22)
- [July 18, 2022](#july-18-22)
]

---
class: inverse, middle
# July 11, 2022

---
class: inverse

.pull-left[

.stress[Mental Stress and Autonomic Reactivity] 
- Result section being completed
- Table and figures complete
- _Share with coauthors_

.stress[HRV and Coronary Vascular Tone]
- Enrolling new patients with continuous ECG recordings 
- Revisitng analyses

.clin[Endocarditis]
- _Outlining_ review paper on arrhythmias and IE 
- Pixel density being _analyzed_ for vegetation chronicity
- IRB pending
]

.pull-right[
.rhythm[AF Registry]
- _Revising_ data dictionary
- Obtaining EGM data from EP stations at UIC and JBVA

.rhythm[Isoproterenol washout and membrane potentials]
- Researching background data
- Evaluating potential test models (animal versus cellular)

.comp[Modeling software]
- _Revising_ R package on model formulas and DAGs

.comp[Circadian rhythms]
- Algorithm for extracting rhythm patterns from continuous data
- Cosinor analysis modification in R
]

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]


---
class: inverse, middle
# July 18, 2022

---
class: inverse, center

```{r}
updates <-
	updates |>
	dplyr::arrange(desc(Effort)) 

tbl_research_updates(updates)
```

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---

## AF Registry

The AF registry is described [here](https://asshah4.github.io/afib). Current issues:

1. Data quality - consistency between reviewers
1. Missingness - decisions on acceptable thresholds
1. Variable selection - echo findings, EP studies, medications, labs, symptoms
1. Management - REDCap, shared excel sheet
1. Adjudication/review - outcomes, clinical follow-up length

.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---
class: inverse, middle
# July 25, 2022

---
class: inverse, center

```{r}
updates <-
	updates |>
	mutate(Status = case_when(
		Project == "MS and CV Mortality" ~ 
			"Team = Amit Shah (*Emory*)  
			Final figures of mortality by subgroups **7/25**  
			Revise discussion on vagal withdrawal **6/26**",
		Project == "AF Registry" ~ 
			"Decide on retrospective/prospective cohort definitions **TBD**",
		Project == "PERT" ~ 
			"Submit IRB for education and clinical data **7/28**",
		Project == "F32/T32/K23" ~ 
			"Aim 1 = Human vagolysis and SCD  
			**Aim 2 = Murine model of vagal withdrawal and arrhythmia**  
			Active research into model development **daily**",
		Project == "Studying" ~ 
			"General cardiology/echo boards in **Fall 2023",
		Project == "UIC Clinic" ~ 
			"Weekly clinic with Dr. Ziccardi",
		Project == "Endocarditis and Pixel Density" ~ 
			"Team = Lee Frazin, Satish Mishra, David Polman
			Revise still images to create standard for processing and SOP **7/30**",
		Project == "Endocarditis and Arrhythmias" ~ 
			"Meet with Atreya Mishra to complete outline **07/28**",
		Project == "Modeling Software" ~ 
			"Forest plots by subgroups in {R}, **drafted**",
		Project == "EGM Signal Processing" ~ 
			"Convert CARTO data into OpenEP **7/31**",
		Project == "HRV and CAD" ~ 
			"Emory Biobank continually recruiting  
			Meet with staff to assess data quality **TBD**",
		Project == "EP Lab" ~ 
			"Discuss with Adam which cases to cover this week  ",
		TRUE ~ Status
	)) |>
	mutate(Effort = case_when(
		Project == "MS and CV Mortality" ~ 0.2,
		Project == "AF Registry" ~ 0.1,
		Project == "PERT" ~ 0.1,
		Project == "F32/T32/K23" ~ 0.2,
		Project == "Studying" ~ 0.03,
		Project == "UIC Clinic" ~ 0.1,
		Project == "Endocarditis and Pixel Density" ~ 0.1,
		Project == "Endocarditis and Arrhythmias" ~ 0.02,
		Project == "Modeling Software" ~ 0.05,
		Project == "EGM Signal Processing" ~ 0.05,
		Project == "HRV and CAD" ~ 0.01,
		Project == "EP Lab" ~ 0.1,
		TRUE ~ Effort
	)) |>
	arrange(desc(Effort))
	
tbl_research_updates(updates)

```


.footnote[
.stress[Stress Epidemiology], .rhythm[Arrhythmia Mechanisms], .clin[Clinical], .comp[Computational]
]

---

## Forest Plots

.pull-left[
```{r}
library(volundr)
x <- rx(Petal.Length ~ X(Sepal.Length) + Petal.Width + S(Species),
				label = list(Petal.Length ~ "Length of Petals",
										 Species ~ "Genus of the Flower"),
				pattern = "direct")
f <- fmls(x, order = 2) 
m <- 
	fit(f, .fit = lm, data = iris, archetype = TRUE) |>
	mdls()

tbl <- tbl_forest(
	object = m,
	y = "Petal.Length",
	x = "Sepal.Length",
	groups = "Species",
	columns = list(beta ~ "Beta Estimate",
								 conf ~ "95% Confidence Interval",
								 n ~ "No."),
	axis = list(lim ~ c(0, 3),
							breaks ~ c(0, 1, 2),
							lab ~ "ß (95% CI)",
							title ~ "Association of Petal Length with Sepal Length")
)
```
]

.pull-right[
```{r, echo = FALSE}
tbl
```
]

